### 2017-08-02 2017mergingmerges/section/01-Introduction.tex
start on final edits

The @group@ operator detects groups of consecutive identical elements and emits a single representative, while @merge@ combines two sorted streams so that the output remains sorted. This example has a few interesting properties. Firstly, the data-access pattern of @merge@ is \emph{value-dependent}, meaning that the order in which @merge@ pulls values from @sIn1@ and @sIn2@ depends on the values themselves: at each step, @merge@ must compare the values from both streams, and choose the stream with the smaller value to pull from.

@@

To implement this program we might write each operator as its own concurrent process, sending stream elements over intra-process channels. Developing this could be easy or hard, depending on the available language features for concurrency. However, the \emph{performance tuning} of such a system, such as using back-pressure to prevent buffers from being overrun, or how to chunk stream data to amortize communication overhead, is invariably a headache.

>>>>

The @group@ operator detects groups of consecutive identical elements and emits a single representative, while @merge@ combines two sorted streams so that the output remains sorted. This example has a few interesting properties. Firstly, the data-access pattern of @merge@ is \emph{value-dependent}, meaning that the order in which @merge@ pulls values from @sIn1@ and @sIn2@ depends on the values themselves.
%: at each step, @merge@ must compare the values from both streams, and choose the stream with the smaller value to pull from.

@@

% To implement this program we might write each operator as its own concurrent process, sending stream elements over inter-process channels.
To implement this we might translate each operator into its own process in a concurrent network. This could be easy or hard, depending on the available language features for concurrency. However, the \emph{performance tuning} of such a system, such as using back-pressure to prevent buffers from being overrun, or chunking stream data to amortize communication overhead, is invariably a headache.

### 2017-08-02 2017mergingmerges/section/XX-Related.tex
start on final edits

In relation to process calculi, synchronised product has been suggested as a method for fusing Kahn process networks together~\cite{fradet2004network}, but does not appear to have been implemented or evaluated. The synchronised product of two processes allows either process to take independent or local steps at any time, but shared actions, such as when both processes communicate on the same channel, must be taken in both processes at the same time. When two processes share multiple channels, synchronised product will fail unless both processes read the channels in exactly the same order. In our system the use of stream buffer variables allows some leeway in when processes must take shared steps.

@@

Synchronous dataflow (SDF; not to be confused with synchronous languages above) is a dataflow graph model of computation where each node has constant, statically known input and output rates. StreamIt~\cite{thies2002streamit} uses synchronous dataflow for scheduling when possible, otherwise falling back to dynamic scheduling~\cite{soule2013dynamic}. Boolean dataflow and integer dataflow~\cite{buck1993scheduling,buck1994static} extend SDF with boolean and integer valued control ports, but only support limited control flow structures.

>>>>

Filter Fusion~\cite{proebsting1996filter} uses a similar algorithm, but only supports processes with at most one input and one output channel.
Each output channel may only be used once.
This means combinators such as @zip@, @unzip@, @append@, and @partition@ cannot be expressed.
Here a consumer and a producer are fused by alternating between the two; when the consumer needs input, control is passed to the producer, and when the producer has output, control is passed to the consumer.
This is sufficient when only one channel is shared, but when multiple channels are shared finer-grained coordination is required.

% This is sufficient for producer-consumer fusion, but two consumers pulling from the same input channel need not follow the same alternation.
% This is sufficient when only one , but when multiple channels are shared between two processes, as may be the case in an intermediate fused result of a larger program, it is not obvious whether simply alternating the two processes is correct.

% The synchronised product of two processes allows either process to take independent or local steps at any time, but shared actions, such as when both processes communicate on the same channel, must be taken in both processes at the same time.
% When two processes share multiple channels, synchronised product will fail unless both processes read the channels in exactly the same order.
% In our system the use of stream buffer variables allows some leeway in when processes must take shared steps.

@@

Synchronous dataflow (SDF; not to be confused with synchronous languages above) is a dataflow graph model of computation where each node has constant, statically known input and output rates. StreamIt~\cite{thies2002streamit} uses synchronous dataflow for scheduling when possible, otherwise falling back to dynamic scheduling~\cite{soule2013dynamic}. Boolean dataflow and integer dataflow~\cite{buck1993scheduling,buck1994static} extend SDF with boolean and integer valued control ports, but only support limited control flow.
